---
title: "The Kids Theming"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The Kids Theming}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.width = 6
)

```

# Overview

Using functions that automatically apply a set of formatting options to plots and tables saves time, allowing us to focus on the analysis and interpretation. Code will also look much cleaner when those ~10 repeated lines of ggplot for each plot are automated away. Importantly, these functions also ensure a polished and consistent appearance across our team, so that outputs look the same irrespective of who generated it.

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(thekidsbiostats)
```

All theming functions host a variety of other options that can further tweak the overall look of plots and tables (including colours).

## The Kids Colours
Colours, consistent with "The Kids" brand guidelines, can be accessed directly via `thekids_colours` list:

```{r echo=F}

base_colours <- names(thekids_palettes$primary)

# Build a tibble with ColourFamily as rows and Palettes as columns
df_colours <- tibble(
  Colour = base_colours,
  Primary = sapply(base_colours, function(x) grep(paste0("^", tolower(x), "$"), names(thekids_colours), value = TRUE)),
  `50%`   = sapply(base_colours, function(x) grep(paste0("^", tolower(x), "_50$"), names(thekids_colours), value = TRUE)),
  `10%`   = sapply(base_colours, function(x) grep(paste0("^", tolower(x), "_10$"), names(thekids_colours), value = TRUE))
)

df_colours |> 
  flextable::flextable() |> 
  align(part='all', j=2:4, align='center') |> 
  padding(part='all', j = 1,  padding.left=6, padding.right = 20) |> 
  padding(part='all', j = 2:4,  padding.left=8, padding.right = 12) |> 
  font(j=2:4, fontname = "Courier New", part = "body") |> 
  bold(part='header', bold=TRUE) |> 
  vline(j=1) |> 
  fontsize(part='header', size=12)
  
```

<br>

The colours can be visualised with the `thekids_showpalette()` function, i.e.

```{r echo=T}
thekids_showpalette()
```


<br><br>

# Example Usage

## `> The Kids theming functions`

The `thekids_theme`, `scale_colour_thekids` and `scale_fill_thekids` functions are useful to apply consistent theming to `ggplot2` visualisations. These use a clean, minimal aesthetic with fonts that align with "The Kids" branding. Here’s an example of a plot before-and-after theming:

```{r echo=TRUE}
ggplot(mtcars, aes(x = mpg, y = wt, colour = factor(cyl))) +
  geom_point(size = 3) +
  labs(x = "Miles per Gallon", y = "Weight", colour = "Cylinders")

```

```{r echo=TRUE}
ggplot(mtcars, aes(x = mpg, y = wt, colour = factor(cyl))) +
  geom_point(size = 3) +
  labs(x = "Miles per Gallon", y = "Weight", colour = "Cylinders") +
  thekids_theme() +
  scale_colour_thekids()

```

<br>

*Miles* better! And as simple as just adding these lines of code: `thekids_theme() + scale_colour_thekids()`.

## `> thekids_table`

`thekids_table` produces tables styled with The Kids branding and is powered by the [flextable](https://cran.r-project.org/web/packages/flextable/index.html) package. By default, this function applies the Barlow font, compact formatting (our preference!), and zebra-striping for readability. Note that these can all be altered/disabled via parameters (see `?thekids_table` documentation).

For example, a raw table output from the mtcars dataset looks like this:

```{r echo=TRUE}
head(mtcars, 5)
```

Now, applying `thekids_table` transforms it into a clean, visually appealing format with branded elements:

```{r echo=TRUE}
head(mtcars, 5) %>%
  thekids_table(colour = "Saffron", font.size = 10)
```

This outputs a compact, zebra-striped table ready to be added to any Word or HTML reports.

The `padding` and `striped` options can be changed if we would prefer more space and to lose the stripes:

```{r echo=TRUE}
head(mtcars, 5) %>%
  thekids_table(colour = "Saffron", 
                font.size = 10, 
                padding = 4, 
                striped = F)
```

