---
title: "The Kids Theming"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The Kids theming}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.width = 6
)
```

# Overview

Using functions that automatically apply a set of formatting options to plots and tables saves time, allowing us to focus on the analysis and interpretation. Code also looks a lot cleaner when those 10 lines of ggplot per plot are automated away. Importantly, these functions also ensure a polished and consistent appearance across our team, so that output looks the same irrespective of who generated it.

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(thekidsbiostats)
```

All theming functions host a variety of other options that can tweak the overall look of plots and tables (including colours).

Colours, consistent with "The Kids" brand guidelines, can be accessed directly via `thekidsbiostats::thekids_colours`:

```{r}
thekidsbiostats::thekids_colours
```


# Example Usage

## `thekids_theme`

The `thekids_theme` function applies consistent theming to ggplot2 visualizations. It uses a clean, minimal aesthetic with fonts and colours that align with "The Kids" branding. Here’s an example of a plot before theming:

```{r echo=TRUE}
ggplot(mtcars, 
       aes(x = mpg, y = wt, col = factor(cyl))) +
  geom_point(size = 3) +
  labs(x = "Miles per Gallon", y = "Weight", col = "Cylinders")
```

```{r}
ggplot(mtcars, 
       aes(x = mpg, y = wt, col = factor(cyl))) +
  geom_point(size = 3) +
  labs(x = "Miles per Gallon", y = "Weight", col = "Cylinders") +
  thekids_theme()
```

## `thekids_table`

`thekids_table` produces tables styled with The Kids branding and is powered by the [flextable](https://cran.r-project.org/web/packages/flextable/index.html) package. This includes by default applying the Barlow font, compact formatting (our preference!), and zebra-striping for readability, though these can all be altered/disabled via parameters.

For example, a raw table output from the mtcars dataset looks like this:

```{r echo=TRUE}
head(mtcars, 5)
```

Now, applying `thekids_table` transforms it into a clean, visually appealing format with branding elements:

```{r echo=TRUE}
head(mtcars, 5) %>%
  thekids_table(colour = "Saffron", font.size = 10)
```

This outputs a compact, zebra-striped table ready for inclusion in Word or HTML reports. However, the padding and striped options can be changed if we would prefer some more space without any stripes:

```{r echo=TRUE}
head(mtcars, 5) %>%
  thekids_table(colour = "Saffron", 
                font.size = 10, 
                padding = 4, 
                striped = F)
```

